---
title: 'Spotify Popularity Prediction'
output:
  html_document: default
  pdf_document: default
---

```{r setup, message = FALSE, warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
library(leaps)
library(boot)
set.seed(743)
```



```{r read file, eval = F}
# Set it as current directory
data_dir <- "C:\Users\Richard\Spotify-Popularity-Predictor"

# List all the CSV files in the directory
csv_files <- list.files(paste0(data_dir, "/data"), pattern = ".*csv", full.names = TRUE)

# Read and combine the CSV files using map function and read_csv
combined_data <- map(csv_files, read_csv, show_col_types = FALSE) %>%
  bind_rows()

saveRDS(combined_data, file = "combined_data.rds", compress = FALSE)
```


```{r reading files again}
combined_data <- readRDS("combined_data.rds")
head(combined_data, n = 15)
```


### Quantifying Data Properties

```{r}
number <- dim(combined_data)
number
```
There are 114000 rows and 20 columns

Each row represents a track on Spotify, with details about the track such as `artist`, `album`, `popularity` and other musical attributes like `tempo` and `key`.

```{r 1c}
unique_genres <- unique(combined_data$track_genre)
number_of_genres <- length(unique_genres)
number_of_genres
```
There are 114 genres in this data set.


First we need to remove the songs with zero popularity.
```{r 2}
filtered_data <- combined_data %>% filter(popularity > 0)
```

```{r}

# Find the top N genres by mean popularity
top_10_genres <- combined_data %>%
  group_by(track_genre) %>%
  summarise(mean_popularity = mean(popularity)) %>%
  top_n(10, mean_popularity)

# Filter the data to include only the top 10 genres
top_genres_data <- combined_data %>%
  filter(track_genre %in% top_10_genres$track_genre)

# Create the boxplot
ggplot(top_genres_data, aes(x = popularity, y = reorder(track_genre, -popularity))) +
  geom_boxplot() +
  xlab("Popularity") +
  ylab("Genre") +
  ggtitle("BoxPlot of Popularity by Top 10 Genres")
```


```{r}
num_rows_removed <- nrow(combined_data) - nrow(filtered_data)
num_rows_removed
```
`16020` Rows of data removed when filtering songs with 0 popularity. 

We filtered out the data with 0 popularity as it could represent missing or skewed data. This could affect our prediction model.

```{r}
colSums(is.na(filtered_data))
```
There are no NA values in the `filtered_data`.

Some inappropriate variables for modelling are `track_id` because this is just a unique identifier and does not help with predictions. `album_name` and `track_name` doesn't really help with predictions either, however, `artists` may help because a song from Taylor Swift for example will be much more popular than a random indie pop band.

`track_genre` is a categorical variable and needs to be encoded for modelling. We can use one-hot encoding to transform these into numerical variables if we want to use it for our model.


Model selection choices:
Target Variable - `popularity` will be the target variable.

```{r}
allyhat <- function(xtrain, ytrain, xtest, lambdas, nvmax = 50) {
  # Number of observations in the training data
  n <- nrow(xtrain)
  
  # Initialise a matrix to store the predicted responses
  yhat <- matrix(nrow = nrow(xtest), ncol = length(lambdas))
  
  # Perform backward subset selection on the training data
  search <- regsubsets(xtrain, ytrain, nvmax = nvmax, method = "back")
  
  # Get a summary of the subset selection result
  summ <- summary(search)
  
  # Loop over each value of lambda
  for (i in 1:length(lambdas)) {
    # Calculate the penalized MSE for models with different numbers of predictors
    penMSE <- n * log(summ$rss) + lambdas[i] * (1:nvmax)
    
    # Find the model with the smallest penalized MSE
    best <- which.min(penMSE)
    
    # Get the coefficients of the best model
    betahat <- coef(search, best)
    
    # Get the predictors in the best model
    xinmodel <- cbind(1, xtest)[, summ$which[best, ]]
    
    # Calculate the predicted responses for the test data
    yhat[, i] <- xinmodel %*% betahat
  }
  
  # Return the matrix of predicted responses
  yhat
}
```


```{r fitting}
# Create a new column with the selected variables
spotify_data <- filtered_data[, c(
  "danceability",
  "energy",
  "loudness",
  "speechiness",
  "acousticness",
  "instrumentalness",
  "liveness",
  "valence",
  "tempo",
  "time_signature",
  "popularity"
)]

X <- as.matrix(filtered_data[, -c(1:5, 20)])

# Extract the response variable
y <- filtered_data$popularity

set.seed(743)

folds <- sample(rep(1:10, length.out = nrow(X)))

# Define a set of lambda values
lambdas <- c(2, 4, 6, 8, 10, 12)

# Init a matrix to store fitted values
fitted <- matrix(nrow = nrow(X), ncol = length(lambdas))

# Perform cross-validation using the allyhat function 
for (k in 1:10) {
  train <- (1:nrow(X))[folds != k]
  test <- (1:nrow(X))[folds == k]
  fitted[test, ] <- allyhat(X[train, ], y[train], X[test, ], lambdas, nvmax = 14)
}

result <- rbind(lambdas, colMeans((y - fitted) ^ 2))
print(result)

opt_lambda = 6
```

seems like $\lambda$ = 2,4,6 has the lowest MSPE  of 344.1721

```{r finding coeff}
search <- regsubsets(X, y, nvmax = 14, method = "backward")
summ <- summary(search)


penMSE <- nrow(X)*log(summ$rss)+ opt_lambda*(1:14)
best <- which.min(penMSE)
betahat <- coef(search, best)
betahat |> 
  as_tibble(rownames = "Variable") |> 
  rename(Coeff = value) |> 
  mutate(Coeff = round(Coeff, 3)) |> 
  knitr::kable()
```
Based on this, we can select the model predictors. It seems like `speechiness`, `instrumentalness`, `valence`, `energy`, `dancebility`and `,explicit` all has a relative big impact on the popularity. 

Our target variable is still `popularity`. 

I used a linear regression model with all two-way interactions. This allows for a quick model and capture the potential relationships between the features without increasing computational complexity too much.

I used a lambda value of 6 at the end, even though 2, 4 and 6 all shared the same MSPE.

However, we haven't checked the assumptions such as linearity, normality of residuals and homoscedascity yet.
## Task 1.2.4

```{r}
# Fit final model using selected predictors
final_model <- lm(popularity ~ speechiness + instrumentalness + valence + energy + danceability + explicit, 
            data = filtered_data)

# Summary of the final model
summary(final_model)
```

```{r}
hist(resid(final_model))
```

The residuals look pretty normally distributed, which means our model is fine.

After fitting the final model with our selected predictors, I ran it through the summary function. All the predictors/coefficients have p-value of <0.05, which means they are all statistically significant. 

However, we have a R-squared value of 0.062, which means our model only explains 6.2% of the variations in `popularity`. This means that  while the model does capture some underlying pattern in the data, the low R-squared sugggests there are other factors that I haven't included in this model that contributes to the popularity.


## Testing the prediction with a random song selection

```{r bootstrapping}
set.seed(743)

# Choose a song from the cleaned dataset
my_song <- filtered_data |> 
  slice(614)

# Display the data for the song using kable()
knitr::kable(my_song)

# Prediction function for bootstrapping
pred_func <- function(data, i){
  fit <- lm(popularity ~ speechiness + instrumentalness + valence + energy + danceability + explicit,
            data = data[i,])
  predict(fit, my_song, interval = "none")
}

# Perform bootstrapping with at least 1000 iterations

boots <- boot(filtered_data, pred_func, R = 1000)


# Calculate 95% confidence interval for the prediction
ci <- boots$t0 + c(-1.96, 1.96) * sd(boots$t)
round(ci, 2)
```

The confidence interval is 41.98 to 42.42


# References
Maharshi Pandya. (2022). 🎹 Spotify Tracks Dataset [Data set]. Kaggle. https://doi.org/10.34740/KAGGLE/DSV/4372070





